---
title: "The recession and student loans"
author: "Roger, Adriana, Lilly, David, Joe"
date: "November 6, 2017"
output: ioslides_presentation
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(shiny)
library(readxl)
library(dplyr)
library(ggplot2)
library(tidyr)

# Loading in data
collegedata <- read_excel("MungingProjectsDataofInterest.xlsx")

# Turning the state column into a factor
collegedata$STABBR <- factor(collegedata$STABBR)

# Renaming columns
independent_variables <- c("Percentage of fist time college students", 
                           "Houlsehold Income of Dependent Students", 
                           "Income of Independent Students", 
                           "Median Student Debt at Graduation",
                           "Percentage of Students that are Dependents"
                           )

colnames(collegedata)[9:13] <- independent_variables
```

## Recession and loans 

According to the Federal Reserve Bank of New York, student loan debt has grown more than auto loan, credit card, and home equity loan debt combined since 2003. 

The goal of this project is to take a look at descriptive statistics to see how things have changed for students since the 2008 economic crisis.

## Interactive Plot

```{r metrics}
# Shiny app
inputPanel(
  selectizeInput("states",
                 "State",
                 levels(collegedata$STABBR),
                 selected = "FL",
                 multiple = TRUE), 
  checkboxGroupInput("variables",
                     "Variables to Plot:",
                     choices = independent_variables,
                     selected = independent_variables[2]))

renderPlot(
    collegedata[collegedata$STABBR==input$states,] %>%
      group_by(Year) %>% 
      select(input$variables) %>%
      mutate_all(median) %>%
      gather(variable, value, -Year) %>%
      ggplot() + geom_line(aes(x=Year, y=value, color=variable))
)
```

## Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## R Output

```{r cars}
summary(cars)
```


